%% This file is for producing a Doctoral Thesis proposal.  It should be fairly
%% self-explanatory.  

\documentclass{article}

\usepackage{xspace}
\usepackage{cite,url}

\begin{document}

%\bibliographystyle{plain}
\pagestyle{empty}
\markboth{{\sc thesis proposal}}{{\sc thesis proposal}}
\def\title{Design and Implementation of a Congestion Control Plane}
\def\author{Akshay Narayan}
\def\addrone{32 Vassar St 32-G982}
\def\addrtwo{Cambridge, MA  02139}

\def\degree{Master of Science}
%% Added \deptname for PhD proposals in other fields.
%% Krishna Sethuraman (1990)
\def\deptname{Electrical Engineering \\ and Computer Science}
\def\laboratory{Computer Science and Artificial Intelligence Laboratory}

\def\submissiondate{\today}
\def\completiondate{February 2018}

\def\supervisor{Professor Hari Balakrishnan}
\def\supertitleone{Associate Professor of Electrical Engineering}
\def\supertitletwo{and Computer Science}

\def\readerone{Professor Mohammad Alizadeh}
\def\readeronetitleone{Professor of Electrical Engineering}
\def\readeronetitletwo{and Computer Science}

\def\readertwo{}
\def\readertwotitleone{}
\def\readertwotitletwo{}

\def\readerthree{}
\def\readerthreetitleone{}
\def\readerthreetitletwo{}


\def\abstract{
With Moore's law ending, the gap between general-purpose processor
speeds and network link rates is widening. This trend has led to new
packet-processing ``datapaths'' in endpoints, including kernel
bypass software and emerging SmartNIC hardware.
In addition, several applications are rolling out their own protocols atop UDP (e.g.,
QUIC, WebRTC, Mosh, etc.), forming new datapaths different from the
traditional kernel TCP stack.
All these datapaths require congestion control,
but they must implement it separately because it is not possible
to reuse the kernel's TCP implementations. This paper proposes
moving congestion control from the datapath into a separate agent.
This agent, which we call the {\em congestion control plane\/} (CCP),
must provide both an expressive congestion control API as well as
a specification for datapath designers to implement and deploy CCP.\@
We propose an API for congestion control, datapath primitives, and a
user-space agent design that uses a batching method to communicate
with the datapath.
Our approach promises to preserve the behavior and performance of in-datapath
implementations while making it significantly easier to implement and deploy new
congestion control algorithms.
}

\newcommand{\cut}[1]{}
%
\newcommand{\paragraphnone}[1]{\vspace{0.075in}\noindent{\bf #1}}
\newcommand{\paragrapha}[1]{\vspace{0.075in}\noindent{\bf #1.}}
\newcommand{\paragraphq}[1]{\vspace{0.075in}\noindent{\bf #1}}
%
\newcommand{\eg}{e.g., }
\newcommand{\etc}{{etc.}\xspace}
\newcommand{\ie}{i.e., }
\newcommand{\ccp}{CCP\xspace}
%
\newcommand{\datapath}{datapath\xspace}
\newcommand{\datapaths}{datapaths\xspace}
\newcommand{\Datapaths}{Datapaths\xspace}
\newcommand{\Datapath}{Datapath\xspace}
%
\frenchspacing
%
\newcommand{\handlers}{callbacks\xspace}
\newcommand{\userspace}{user-space\xspace}
\newcommand{\kernelspace}{kernel-space\xspace}
%
\newcommand{\controller}{agent\xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% You Should Not Need To Modify Anything Below Here %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Cover Page - Author signs %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\Large \bf 
   Massachusetts Institute of Technology
\\ Department of \deptname \\}
\vspace{.25in}
{\Large \bf
   Proposal for Thesis Research in Partial Fulfillment
\\ of the Requirements for the Degree of
\\ \degree \\}
\end{center}

\vspace{.5in}

\def\sig{{\small \sc (Signature of Author)}}

\begin{tabular}{rlc}
   {\small \sc Title:}                       & \multicolumn{2}{l}{\title}
\\ {\small \sc Submitted by:}
                            & \author  & \\
                            & \addrone & \\
                            & \addrtwo & \\ \cline{3-3}
			    &	       & \makebox[2in][c]{\sig}
\\ {\small \sc Date of Submission:}          & \multicolumn{2}{l}{\submissiondate}
\\ {\small \sc Expected Date of Completion:} & \multicolumn{2}{l}{\completiondate}
\\ {\small \sc Laboratory:}                  & \multicolumn{2}{l}{\laboratory}
\end{tabular}


\vspace{.75in}
{\bf \sc Brief Statement of the Problem:}

\abstract

         %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage %%% Supervision Agreement %%%
         %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{flushright}
   Massachusetts Institute of Technology
\\ Department of \deptname
\\ Cambridge, Massachusetts 02139
\end{flushright}

\underline{\bf Thesis Supervision Agreement}

\vspace{.25in}
\begin{tabular}{rl}
   {\small \sc To:}   & Department Graduate Committee
\\ {\small \sc From:} & \supervisor
\end{tabular}

\vspace{.25in}
The program outlined in the proposal:

\vspace{.25in}
\begin{tabular}{rl}
   {\small \sc Title:}  & \title
\\ {\small \sc Author:} & \author
\\ {\small \sc Date:}   & \submissiondate
\end{tabular}

\vspace{.25in}
is adequate for a \degree{} thesis.
%I believe that appropriate readers for this thesis would be:
%
%\vspace{.25in}
%\begin{tabular}{rl}
%   {\small \sc Reader 1:} & \readerone
%\\ {\small \sc Reader 2:} & \readertwo
%%\\ {\small \sc Reader 3:} & \readerthree
%\end{tabular}

\vspace{.25in}
Facilities and support for the research outlined in the proposal are available.
I am willing to supervise the thesis and evaluate the thesis report.

\vspace{.25in}
\begin{tabular}{crc}
  \hspace{2in} & {\sc Signed:} & \\ \cline{3-3}
               &               & {\small \sc \supertitleone} \\
               &               & {\small \sc \supertitletwo} \\
               &               &                             \\
               & {\sc Date:}   & \\ \cline{3-3}
\end{tabular}

\vspace{0in plus 1fill}

Comments: \\
\begin{tabular}{c}
  \hspace{6.25in} \\
  \mbox{} \\ \cline{1-1} \mbox{} \\
  \mbox{} \\ \cline{1-1} \mbox{} \\
  \mbox{} \\ \cline{1-1} \mbox{} \\
%  \mbox{} \\ \cline{1-1} \mbox{} \\
%  \mbox{} \\ \cline{1-1} \mbox{} \\
%  \mbox{} \\ \cline{1-1} \mbox{} \\
\end{tabular}

\newpage

\input{ccp-hotnets-intro}
\input{procedure}

\bibliographystyle{abbrv}
\begin{small}
\bibliography{thesis}
\end{small}

\end{document}
